---
title: "Cleanup"
output: html_document
---

```{r setup, include=TRUE}
library(haven)
library(tidyverse)

# Read federal riding census data
riding <- read_csv('98-401-X2016045_English_CSV_data.csv')

# Filter off provincial/national aggregates
riding <- riding %>%
  subset(GEO_LEVEL == 2)
```

```{r setup1, include=TRUE}
# Get human-readable names for columns
prompt <- riding[1:2247,9]

riding_male <- riding %>%
  select(
    GEO_NAME,
    colnames(riding)[10],
    colnames(riding)[13]
    )

riding_female <- riding %>%
  select(
    GEO_NAME,
    colnames(riding)[10],
    colnames(riding)[14]
  )

riding_overall <- riding %>%
  select(
    GEO_NAME,
    colnames(riding)[10],
    colnames(riding)[12]
  )
```

```{r 3, include=TRUE}
names(riding_male)[1] <- 'Riding'
names(riding_male)[2] <- 'Code'
names(riding_male)[3] <- 'Val'

names(riding_female)[1] <- 'Riding'
names(riding_female)[2] <- 'Code'
names(riding_female)[3] <- 'Val'

names(riding_overall)[1] <- 'Riding'
names(riding_overall)[2] <- 'Code'
names(riding_overall)[3] <- 'Val'

riding_male <- riding_male %>%
  pivot_wider(names_from = Code, values_from = Val)

riding_female <- riding_female %>%
  pivot_wider(names_from = Code, values_from = Val)

riding_overall <- riding_overall %>%
  pivot_wider(names_from = Code, values_from = Val)

riding_male$sex <- c('male')
riding_female$sex <- c('female')
```

```{r 4, include=TRUE}
riding_name <- 1:1
age_cols <- 9:33
income_cols <- 760:779
home_ownership_cols <- 1618:1620
language_cols <- 101:111
education_cols <- 1684:1697
ethnicity_cols <- 1325:1410
home_cols <- 1676
immigrants_cols <- 1141:1142
sex <- 2249
employment_cols <- 1867:1870

riding_male <- riding_male %>%
  select(riding_name,
         age_cols,
         income_cols,
         home_ownership_cols,
         language_cols,
         education_cols,
         home_cols,
         immigrants_cols,
         sex,
         employment_cols,
         ethnicity_cols)

riding_female <- riding_female %>%
  select(riding_name,
         age_cols,
         income_cols,
         home_ownership_cols,
         language_cols,
         education_cols,
         home_cols,
         immigrants_cols,
         sex,
         employment_cols,
         ethnicity_cols)

riding_overall <- riding_overall %>%
  select(riding_name,
         age_cols,
         income_cols,
         home_ownership_cols,
         language_cols,
         education_cols,
         home_cols,
         immigrants_cols,
         employment_cols,
         ethnicity_cols)

riding_data <- rbind(riding_male, riding_female)
'%ni%' <- Negate('%in%')
```

age_cols <- 8:32
-------------------------------------------
           colname
15-24,   14, 15
25-34,   16, 17
35 - 44, 18, 19
45 - 54, 20, 21
55 - 64, 22, 23
65 - 74, 25, 26
75 - 84, 27, 28
85+      29 

```{r Age Categories, include=TRUE}
# these are the column names NOT INDICES of corrsponding age brackets
adj_col_names = c("14", "15", "16", "17", "18", "19", "20",
                  "21", "22", "23", "25", "26", "27", "28",
                  "29")

# turning the necessary columns numeric 
for (var in adj_col_names){
  riding_data[[var]] = as.numeric(riding_data[[var]])
}

# dataset containing riding population by age category
Age_Categories = riding_data %>% transmute(
   "Age: 15-24" = rowSums(.[8:9]),
   "Age: 25-34" = rowSums(.[10:11]),
   "Age: 35-44" = rowSums(.[12:13]),
   "Age: 45-54" = rowSums(.[14:15]),
   "Age: 55-64" = rowSums(.[16:17]),
   "Age: 65-74" = rowSums(.[19:20]),
   "Age: 75-84" = rowSums(.[21:22]),
   "Age: 85+"     = rowSums(.[23:23])
)
Age_Categories$Riding = riding_data$Riding


# Sex Adjust
cont = c()
for (i in 1:nrow(Age_Categories)){
  if (i <= 338){
    cont = c(cont, "Male")
  } else {
    cont = c(cont, "Female")
  }
}
cont = as.factor(cont)
Age_Categories$Sex = cont

# Making Riding first column
Age_Categories = Age_Categories[,c(9,10,1,2,3,4,5,6,7,8)]



# Pivor Wider
age_prop <- Age_Categories %>%
  pivot_longer(cols=3:10, names_to="Age Bracket", values_to="n") %>%
  group_by(Riding, Sex) %>%
  mutate(freq = n / sum(n))
```

```{r Home Ownership Categories, include=TRUE}
# IMPORTANT: this variable is not differentiated into sexes
# column names 1618, 1619 refers to Owners and Renters resp.
home_ownership_categories = data.frame(
  Riding = riding_overall$Riding,
  Owner = as.numeric(riding_overall$`1618`),
  Renter = as.numeric(riding_overall$`1619`)
)

# Pivor Wider
home_ownership_prop <- home_ownership_categories %>%
  pivot_longer(cols=2:3, names_to="Status", values_to="n") %>%
  group_by(Riding) %>%
  mutate(freq = n / sum(n))
```


Language Categories: Might need to adjust 

-> Language category by knowledge: 100%
English only: 101
French only: 102
English and French: 103
Neither English nor French: 104

-> Language category by first official spoken: 100%
English: 106
French: 107
English and French: 108	
Neither English nor French: 109	
Official language minority (number): 110	

```{r Language Categories, include=TRUE}
language_categories = data.frame(
  Riding = riding_data$Riding,
  English = riding_data$`106`,
  French = riding_data$`107`,
  English_and_French = riding_data$`108`,
  Neither = riding_data$`109`
)

for (var in colnames(language_categories)){
  if (var != "Riding"){
    language_categories[[var]] = as.numeric(language_categories[[var]])
  }
}
```

```{r Language2, include=TRUE}
library(dplyr)
# Sex Adjust
cont = c()
for (i in 1:nrow(language_categories)){
  if (i <= nrow(language_categories)/2){
    cont = c(cont, "Male")
  } else {
    cont = c(cont, "Female")
  }
}
cont = as.factor(cont)
language_categories$Sex = cont

# Making Riding first column
language_categories = language_categories[,c(1,6,2,3,4,5)]

# Pivor Wider
language_prop <- language_categories %>%
  pivot_longer(cols=3:6, names_to="Official Language Spoken", values_to="n") %>%
  group_by(Riding, Sex) %>%
  mutate(freq = n / sum(n))
```

Canadian -> 1346
English -> 1357
Scottish -> 1360
French -> 1367
Irish -> 1358
German -> 1375
Chinese -> 1326
Italian -> 1409
First Nations -> 1340 
Indian -> 1325

```{r Ethn Categories, include=TRUE}
Ethnic_Categories = data.frame(
  Riding = riding_data$Riding,
  Canadian = riding_data$`1346`,
  English = riding_data$`1357`,
  Scottish = riding_data$`1360`,
  French = riding_data$`1367`,
  Irish = riding_data$`1358`
)

for (var in colnames(Ethnic_Categories)){
  if (var != "Riding"){
    Ethnic_Categories[[var]] = as.numeric(Ethnic_Categories[[var]])
  }
}
```

```{r Ethnic 2, include=TRUE}
library(dplyr)
# Sex Adjust
cont = c()
for (i in 1:nrow(Ethnic_Categories)){
  if (i <= nrow(Ethnic_Categories)/2){
    cont = c(cont, "Male")
  } else {
    cont = c(cont, "Female")
  }
}
cont = as.factor(cont)
Ethnic_Categories$Sex = cont

# Making Riding first column
Ethnic_Categories = Ethnic_Categories[,c(1,7,2,3,4,5,6)]

# Pivor Wider
Ethnic_prop <- Ethnic_Categories %>%
  pivot_longer(cols=3:7, names_to="Ethnic Origins", values_to="n") %>%
  group_by(Riding, Sex) %>%
  mutate(freq = n / sum(n))
```




```{r Write to csv, include=TRUE}
write.csv(Ethnic_prop,"Ethnicity.csv")
write.csv(language_prop,"Language.csv")
write.csv(home_ownership_prop,"Homeownership.csv")
write.csv(age_prop,"Ages.csv")
```








